---
title: "The ExpressionSetDF class"
author: "Alex Sanchez"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc : TRUE
vignette: >
  %\VignetteIndexEntry{The ExpressionSetDF class}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



# Introduction

The ExpressionSetDF class has been created to facilitate the combined use of data and metadata for both, samples and variables, in the same style as can be done with Bioconductor `ExpressionSet` or `SummarizedExperiment` classes.

The main difference is that the Slot "Exprs" (we have kept the name) is a data frame, thus allowing to have variable of distinct types.


# Creating an `expressionSetDF` object

```{r}
require(expressionSetDF)
datos <- data.frame(A = c(1, 2, 3, 7), B = c(4, 5, 6, 0),
            C=c("C", "C", "T", "T"))
sampleNames <- c("S1", "S2", "S3", "S4")
varNames <- c("G1", "G2", "Group")
rownames(datos) <- sampleNames
colnames(datos) <- varNames
phenoDat <-  data.frame(sampleGrp = c(rep("H", 2), rep("A", 2)))
rownames(phenoDat) <- rownames(datos)
featureDat <- data.frame(varGrp= c(rep("c", 2), "n"))
rownames(featureDat) <- colnames(datos)
es_df0 <- expressionSetDF(exprs = datos,
                       phenoData = NULL,
                       featureData = NULL)
show(es_df0)
es_df <- expressionSetDF(exprs = datos,
                      phenoData = phenoDat,
                      featureData = featureDat)
show(es_df)
```


# Accessing the slots

Methods, `exprs.eDF`, `phenoData.eDF` and `featureData.eDF` provide access to the corresponding slots so that they don't need to be accessed directly:

```{r}
exprs.eDF(es_df)
phenoData.eDF(es_df)
featureData.eDF(es_df)
```

# Subsetting

A subsetting method has been defined through the standard brackets system:

```{r}
(es_colset <- es_df[1:2,])
(es_rowset <- es_df[,1:2])
```


## Reordering the slots

Subsetting can also be used to re-order the data.

```{r}

revOrder<- order(rownames(exprs.eDF
                          (es_df)), decreasing=TRUE)
sorted_es <- es_df[revOrder,]
sorted_es
```

